A Speculative Approach to Parallelization in Particle Swarm Optimization

### 2011-02-09 honors_thesis/si_paper/paper.tex
Added a couple of paragraphs discussing some related work

Thus previous work in parallelizing PSO has consisted entirely of increasing
the swarm size and adapting the topology to be better suited to parallel
computation.

@@

We also do not report absolute time used by each algorithm, as specific time
values are heavily implementation-dependent.  Using time steps, where at every
time step each processor evaluates the function once, is an abstraction from
individual implementations that still conveys all of the information necessary
to compare algorithms.  In this paper we focus on the effects of algorithm
choices on parallel PSO performance.  As there are inexpensive functions with
complex landscapes and expensive functions with simple landscapes, the behavior
of PSO with respect to the function is the main issue.  For this purpose,
benchmark functions are a useful and efficient tool for understanding the
effects of PSO with expensive objective functions even though the benchmark
functions are themselves inexpensive.  For example, a plot of performance with
respect to iterations for a smooth unimodel function with five minutes per
evaluation would be similar to that of Sphere.

@@

>>>>

Thus previous work in parallelizing PSO, apart from implementation details, has
consisted entirely of increasing the swarm size and adapting the topology to be
better suited to parallel computation.

With regard to increasing the swarm size in PSO, some recent work has suggested
that increasing the swarm size through the course of the optimization process
provides better results than having a set swarm
size~\citep{hsieh-2009-efficient-population-utilization-for-pso,
montes-de-oca-2010-incremental-social-learning-pso}.  However, these results
are based on total number of function evaluations, which, when running in
parallel on expensive functions, is less important than total number of
iterations.  Other work focusing on parallelization has shown that when extra
processors are available they should be used, as performance increases with
swarm size when measuring in terms of number of
iterations~\cite{mcnabb-2009-large-particle-swarms,
jin-2005-pso-antenna-designs}.  If the swarm size were varied throughout the
course of the optimization process, some processors would be sitting idle at
most iterations.

@@

We also do not report absolute time used by each algorithm, as specific time
values are heavily implementation-dependent.  Using time steps, where at every
time step each processor evaluates the function once, is an abstraction from
individual implementations that still conveys all of the information necessary
to compare algorithms.  In this paper we focus on the effects of algorithm
choices on parallel PSO performance.  As there are inexpensive functions with
complex landscapes and expensive functions with simple landscapes, the behavior
of PSO with respect to the function is the main issue.  For this purpose,
benchmark functions are a useful and efficient tool for understanding the
effects of PSO with expensive objective functions even though the benchmark
functions are themselves inexpensive.  For example, a plot of performance with
respect to iterations for a smooth unimodal function with five minutes per
evaluation would be similar to that of Sphere.

@@

We mentioned related work showing that increasing the swarm size throughout the
course of the algorithm could provide improved performance over a fixed swarm
size in serial PSO~\citep{montes-de-oca-2010-incremental-social-learning-pso}.
If this method were extended to parallel PSO, most processors would be idle in
the first few iterations, while more would be utilized at the end.  During
iterations where there are many un-utilized processors, a natural use of them
would be speculative evaluation, performing two or more of those iterations at
a time.  Thus our methods have the potential to further improve the parallel
performance of recently proposed variations of PSO.
